Emulation of high multiplicity matrix elements for electron-positron and hadron-hadron colliders

### 2022-10-10 mc.tex
add references (MC)

    From this description, it is clear that the number
    of accepted events will generally be lower than
    the number of total events generated. This is encapsulated
    in the unweighting efficiency
    \begin{equation}\label{eqn:unweighting_efficiency}
        \varepsilon = \dfrac{N_{\mathrm{accepted}}}{N_{\mathrm{total}}} \, ,
    \end{equation}
    whose inverse is the average number of trial
    events required to obtain a single unweighted event.
    Unweighting efficiency for high-multiplicity processes
    are often well below 1\% \cite{Hoche:2019flt,Gao:2020zvv} and development
    of methods to improve the efficiency have been
    explored \cite{Jadach:1999sf,Jadach:2002kn}.

@@

    As seen in (\ref{eqn:matrix_element}),
    the exact matrix element is approximated by a perturbative
    series in a coupling constant. The calculations of
    tree-level (cite here) and one-loop (cite here)
    matrix elements are now largely fully automated
    in matrix element generators with a common generci
    interface, the Binoth Les Houches Accord (cite here).
    For more information on the machinery developed to
    compute nearly arbitrary matrix elements, see Reference (cite here).
    NNLO and N$^{3}$LO QCD corrections have been computed
    for select processes but have yet to be reach the
    level of automation of NLO.
    For a review of recent advancements see References (cite here).

    Matrix element calculations make up a large
    portion of the total event generation time (cite here).
    Reducing the time spent in this stage of the event
    generation therefore represents a sizeable increase
    in the number of events generated given a fixed computing
    budget. This thesis presents a novel technique to
    accelerate matrix element calculations by building
    accurate emulation models from modern machine learning
    techniques.

@@

    Parton showers are implemented as algorithms that
    iteratively emit radiation (gluons and quarks in QCD
    parton showers) described by splitting functions.
    By repeating this procedure of emissions, parton
    showers produce a cascade of partons that naturally
    evolves the energy scale from the hard process to
    an infrared scale where non-perturbative effects
    begin to set in. In doing so parton showers account
    for potentially large logarithms that arise in
    these kinematic regimes at all orders in the coupling
    parameter, this is known as resummation.

    In event generators, fixed-order matrix elements
    and parton showers have to be utilised together to describe
    the many different observables at experiments. In order
    to combine the two stages of event generation in a consistent
    manner there are matching (cite here) and merging (cite here)
    schemes.

    \subsection*{Hadronisation}\label{sec:hadronisation}
    The formation of hadrons in the low energy regime
    of an event generator follow non-perturbative models.
    The two main classes of models used to simulate
    hadronisation are the cluster models (cite here)
    and string models (cite here).
    These models share a similar concept of grouping together
    quarks and antiquarks to ensure only hadrons remain
    in the final-state. For more discussion on hadronisation
    see References (cite here).

@@

    The most widely adopted approach to estimate the
    theoretical uncertainties associated with these
    missing higher order terms is to carry out a seven-point
    scale variation. Typically in a calculation we take
    $\mu = \mu_{F} = \mu_{R}$ where $\mu$ depends on the specifics
    of the process of interest. Taking the value of the
    partonic cross-section at this scale as the central value,
    the seven variations correspond to multiplying $\mu_{F}$
    and $\mu_{R}$ by the factors
    \begin{equation}\label{eqn:scale_variations}
        \dfrac{\mu_{i}}{\mu} = \left\{(1, 1), (2,2),(0.5,0.5),(2,1),(1,2),(1,0.5),(0.5,1)\right\} \, .
    \end{equation}
    The estimate of the uncertainty is then taken
    to be the half-width of the maximum interval.

>>>>

    From this description, it is clear that the number
    of accepted events will generally be lower than
    the number of total events generated. This is encapsulated
    in the unweighting efficiency
    \begin{equation}\label{eqn:unweighting_efficiency}
        \varepsilon = \dfrac{N_{\mathrm{accepted}}}{N_{\mathrm{total}}} \, ,
    \end{equation}
    whose inverse is the average number of trial
    events required to obtain a single unweighted event.
    Unweighting efficiency for high-multiplicity processes
    are often well below 1\% \cite{Hoche:2019flt,Gao:2020zvv} and development
    of methods to improve the efficiency have been
    explored \cite{Jadach:1999sf,Jadach:2002kn}.
    These low efficiencies lead to most of the
    yearly CPU budgets of LHC experiments to be spent on
    unweighting events \cite{HSFPhysicsEventGeneratorWG:2020gxw}.

@@

    As seen in (\ref{eqn:matrix_element}),
    the exact matrix element is approximated by a perturbative
    series in a coupling constant. The calculations of
    tree-level \cite{Mangano:2002ea,Krauss:2001iv,Gleisberg:2008fv,Cafarella:2007pc,Alwall:2011uj,Kilian:2007gr}
    and one-loop \cite{Berger:2008sj,Cullen:2011ac,Cullen:2014yla,Bevilacqua:2011xh,Hirschi:2011pa,Alwall:2014hca,Frederix:2018nkq,Campbell:1999ah,Campbell:2021vlt,Badger:2012pg,Cascioli:2011va,Buccioni:2019sur,Actis:2016mpe,Denner:2017wsf}
    matrix elements are now largely fully automated
    in NLO QCD and electroweak corrections.
    Many of the modern matrix element generators are supplied
    with a common generic interface to event generators,
    the Binoth Les Houches Accord \cite{Binoth:2010xt,Alioli:2013nda}.
    For more information on the machinery developed to
    compute nearly arbitrary matrix elements at tree-level and
    one-loop level, see for instance Reference \cite{Campbell:2017hsr}.
    NNLO and N$^{3}$LO QCD corrections have been computed
    for select processes but have yet to be reach the
    level of automation of NLO.
    For a review of recent advancements see References \cite{Caola:2022ayt,Campbell:2022qmc}.

    Matrix element calculations make up a large
    portion of the total event generation time \cite{Bothmann:2022thx}.
    Reducing the time spent in this stage of the event
    generation therefore represents a sizeable increase
    in the number of events generated given a fixed computing
    budget \cite{HSFPhysicsEventGeneratorWG:2020gxw}.
    This thesis presents a novel technique to
    accelerate matrix element calculations by building
    accurate emulation models from modern machine learning
    techniques.

@@

    Parton showers are implemented as algorithms that
    iteratively emit radiation (gluons and quarks in QCD
    parton showers) described by splitting functions.
    By repeating this procedure of emissions, parton
    showers produce a cascade of partons that naturally
    evolves the energy scale from the hard process to
    an infrared scale where non-perturbative effects
    begin to set in. In doing so parton showers account
    for potentially large logarithms that arise in
    these kinematic regimes at all orders in the coupling
    parameter, this is known as resummation \cite{Buckley:2011ms}.

    In event generators, fixed-order matrix elements
    and parton showers have to be utilised together to describe
    the many different observables at experiments. In order
    to combine the two stages of event generation in a consistent
    manner there are matching \cite{Frixione:2002ik,Frixione:2007vw,Jadach:2015mza}
    and merging \cite{Catani:2001cc,Lonnblad:2011xx}
    schemes.

    \subsection*{Hadronisation}\label{sec:hadronisation}
    The formation of hadrons in the low energy regime
    of an event generator follow non-perturbative models.
    The two main classes of models used to simulate
    hadronisation are the cluster models \cite{Webber:1983if,Winter:2003tt}
    and string models \cite{Andersson:1983ia}.
    These models share a similar concept of grouping together
    quarks and antiquarks to ensure only hadrons remain
    in the final-state. For more discussion on hadronisation
    see for example Reference \cite{Webber:1999ui}.

@@

    The most widely adopted approach to estimate the
    theoretical uncertainties associated with these
    missing higher order terms is to carry out a seven-point
    scale variation. Typically in a calculation we take
    $\mu = \mu_{F} = \mu_{R}$ where $\mu$ depends on the specifics
    of the process of interest. Taking the value of the
    partonic cross-section at this scale as the central value,
    the seven variations correspond to multiplying $\mu_{F}$
    and $\mu_{R}$ by the factors 
    \begin{equation}\label{eqn:scale_variations}
        \dfrac{\mu_{i}}{\mu} = \left\{(1, 1), (2,2),(0.5,0.5),(2,1),(1,2),(1,0.5),(0.5,1)\right\} \, ,
    \end{equation}
    and re-evaluating the
    cross-section at these scales.
    The estimate of the uncertainty is then taken
    to be the half-width of the maximum interval.
